---
title: "Sample Integration Report"
author: "CCBR Pipeliner"
date: '`r format(Sys.time(), "%a %b %d %Y - %X")`'
output:
  html_document: default
params:
  gid: "text"
  mergedObj: "mergedObject.rds"
  integratedObj: "integratedObject.rds"
  rpcaObj: "rpcaObject.rds"
  harmonyGroupObj: "harmonyGroupObject.rds"
  harmonySampleObj: "harmonySampleObject.rds"
  resolution: "csvListResolutionString"
  citeseq: "text"
  annot: "text"
  imageDir: "path/to/images"
geometry:
  margins:0.25in
---

```{r, handle_pkg, message=FALSE}
# set library dir, load this and remove any other dirs to avoid confusion
# between personally created pkgs and the pipeline package
## saving old path "/Library/Frameworks/R.framework/Versions/4.2-arm64/Resources/library"
print(paste0("Using the lib.loc location: ",Rlib_dir))
assign(".lib.loc", Rlib_dir, envir = environment(.libPaths))
.libPaths()

# read in package info
pkg_df=read.csv(Rpkg_config)
pkg_df=subset(pkg_df,int=="Y")
pkg_df

# for each package check installation, if present then load library
for (rowid in rownames(pkg_df)){
  pkg=pkg_df[rowid,"package"]
  source=pkg_df[rowid,"source"]
  version=pkg_df[rowid,"version"]
  gh_name=pkg_df[rowid,"gh_name"]

  need_install <- pkg[!(pkg %in% installed.packages()[,"Package"])]
  if (length(need_install)!=0){
    print(paste0("Installing: ", pkg))
    if (source=="bc") BiocManager::install(pkg)
    if (source=="cr") install.packages(pkg,version=version,repos = "http://cran.us.r-project.org",
                                       local = FALSE)
    if (source=="gh") remotes::install_github(gh_name,version=version,local = FALSE)
  }

  print(paste0("Loading: ",pkg))
  invisible(lapply(pkg, library, character.only = TRUE))
}

# source functions
source(scRNA_functions)

# additional options
tinytex::install_tinytex(force = TRUE)
```

```{r libraryImport, message=FALSE, warning=FALSE}
path = dirname(params$imageDir)
require("knitr")
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
knitr::opts_knit$set(
  root.dir=path
)

.libPaths("/data/CCBR_Pipeliner/db/PipeDB/")
.libPaths("/data/CCBR_Pipeliner/db/PipeDB/scrna4.2Rlibs")

#Visualization libraries
library(magick)
library(ggplot2)
library(cowplot)
library(scales)
library(RColorBrewer)
library(ggpubr)

#Bioconductor dependencies
library(BiocGenerics)
library(Biobase)
library(farver)
library(S4Vectors)
library(AnnotationDbi)
# library(modes)
library(dplyr)
library(Matrix)
library(reshape2)
library(tools)
library(stringr)

#statistics libraries
# library(Routliers)
library(VennDiagram)
library(cluster)
library(DT)

#scRNASeq libraries
library(Seurat)
library(scRNAseq)
library(SingleCellExperiment)
library(SingleR)
library(scater)
library(DoubletFinder)
# library(URD)
library(celldex)
```

```{r dataLoad, message=FALSE, warning=FALSE}
# set up params
gid = params$gid
mergedObj=params$mergedObj
integratedObj=params$integratedObj
rpcaObj=params$rpcaObj
harmonyGroup=params$harmonyGroupObj
harmonySample=params$harmonySampleObj

resolution=as.numeric(strsplit(params$resolution_list,",")[[1]])
# Rlib_dir=params$Rlib_dir
# Rpkg_config=params$Rpkg_config
# scRNA_functions = params$scRNA_functions
# 

#"contrasts" may be the same as "group"
#citeseq: "text"
  annot: "text"

# read in SO
so_merged = readRDS(mergedObj)
so_rpca = readRDS(rpcaObj)
so_harmonyGroup = readRDS(harmonyGroup)
so_harmonySample = readRDS(harmonySample)
so_integrated = readRDS(integratedObj)

# identify object list
objList = c("merged","integrated","harmonyGroup","harmonySample","rpca")
```

### Batch correction selection

#### Samples before and after batch correction:
```{r sampleBatchCorrection, fig.width=15, fig.height=15}
for(i in 1:length(objList)){
  obj = OBJECT_SELECT(objList[i])
  title = NAME_SELECT(objList[i])
  sampleImage = DimPlot(obj,group.by="Sample",pt.size = 0.01) + 
    ggtitle("",paste0(title,", colored by Sample"))+
    theme(legend.position="bottom",
          legend.text=element_text(size=12))+
    guides(colour=guide_legend(ncol=6,override.aes = list(size=3)))
  assign(x=paste0("s",i),value=sampleImage)
}
ggarrange(s1,s2,s3,s4,s5,ncol=2,nrow=3,common.legend=T,legend="bottom")
```

```{r groupBatchCorrection, fig.width=15, fig.height=15}
for(i in 1:length(objList)){
  obj = OBJECT_SELECT(objList[i])
  title = NAME_SELECT(objList[i])
  sampleImage = DimPlot(obj,group.by="groups",raster=F, pt.size= 0.001) + 
    ggtitle("",paste0(title,", colored by Groups"))+
    theme(legend.position="bottom",legend.text=element_text(size=12))+
    guides(colour=guide_legend(override.aes = list(size=3)))
  assign(x=paste0("g",i),value=sampleImage)
}
ggarrange(g1,g2,g3,g4,g5,ncol=2,nrow=3,common.legend=T,legend="bottom")
```

#### Cell type annotations before and after batch corrrection
```{r annotationBatchCorrection, fig.width=15, fig.height=15}
for(i in 1:length(objList)){
  obj = OBJECT_SELECT(objList[i])
  title = NAME_SELECT(objList[i])
  sampleImage = DimPlot(obj,group.by="annot") + 
    ggtitle("",paste0(title,", colored by primary annotation"))+
    theme(legend.position="bottom",legend.text=element_text(size=12))+
    guides(colour=guide_legend(ncol=6,override.aes = list(size=3)))
  assign(x=paste0("ct",i),value=sampleImage)
}
ggarrange(ct1,ct2,ct3,ct4,ct5,ncol=2,nrow=3,common.legend=T,legend="bottom")

datatable(as.data.frame.matrix(cbind(table(merged$annot,
                                           merged$Sample),table(merged$annot))),
          colnames=c(unique(merged$Sample),"Total"))
```

### Clustering and silhouette scores before and after batch correction {.tabset}

```{r silhouetteImages,  echo=FALSE, include=FALSE}
# create matrix
resSil_merged =matrix(ncol=2,nrow=length(resString))
resSil_merged[]<-0
resSil_integrated=resSil_merged
resSil_rpca=resSil_merged
resSil_harmonyGroup=resSil_merged
resSil_harmonySample=resSil_merged

# Initial loop through integration methods
for(k in 1:length(objList)){
  obj = OBJECT_SELECT(objList[k])

  i=1
  #Initialize silhouette  scores matrix
  resSil_mat = matrix(ncol=2,nrow=length(resolution))
  resSil_mat[,1]<- as.numeric(resolution)
  resSil_mat[,2]<-0

  #Skip all calculations for objects greater than 90k cells
  if(ncol(obj)>90000){
    for (res in resolution){
      fileName = paste0(params$imageDir,"/silhouetteResolution_",res,"_",objList[k],".pdf")
      pdf(fileName)
      plot.new()
      text(0.5,0.5,"No silhouette available for datasets with more than 90,000 cells")
      dev.off()    
    }
  } else{
    if(length(grep("^harmony",objList[k]))){
      dist = dist(Embeddings(obj,reduction='harmony')[,1:30])
    } else{
      dist = dist(Embeddings(obj,reduction='pca')[,1:30])
    }

    for (res in resolution){
      #convert resolution to numeric value
      resMod=as.numeric(res)   
      #SLM name ident determination, based on integration method
      slm_name = grep("^SLM",colnames(obj@meta.data),value=T)
      if (objList[k] %in% c("integrated","rpca")){
        slm_name=gsub("\\..*","",grep("int",slm_name,value=T)[1])
      }else{
        slm_name=gsub("\\..*","",grep("SCT",slm_name,value=T)[1])
      }
      #create file for resolution and integration method
      fileName = paste0(params$imageDir,"/silhouetteResolution_",res,"_",objList[k],".pdf")
      pdf(fileName)

      slm_name = paste0(slm_name,".",resMod)

      if ((slm_name %in% colnames(obj@meta.data))==FALSE){
        obj = FindClusters(obj, algorithm =3, resolution = resMod)
        obj[[slm_name]] = Idents(obj)
        rdsFile = params[[grep(objList[k],names(params),value=T)]]
        saveRDS(obj,file=rdsFile)
        message(paste(objList[k],
                      "object has been overwritten to include clustering at resolution",
                      res))
      }

      #Calculate silhouette scores for resolutions with more than one cluster
      Idents(obj)=slm_name
      clusters=Idents(obj)

      if(length(unique(clusters))==1){ #edge case of failed sample, only one cluster
        plot.new()
        text(0.5,0.5,"No silhouette available for single cluster samples")
        dev.off()
      } else{
        sil=silhouette(as.numeric(as.character(clusters)),dist=dist)
        palette=alpha(colour=hue_pal()(length(unique(Idents(obj)))),alpha=0.7)
        
        #Export silhouette plot to file
        print(plot(sil, col=palette[as.factor(clusters[order(as.numeric(as.character(clusters)),
                                                             decreasing=F)])],
        main=paste0("Silhouette plot of clustering resolution ", res), lty=2,
        sub=paste("Average silhouette width:",format(round(mean(sil[,3]), 4), nsmall = 4))))
        abline(v=mean(sil[,3]), col="red4", lty=2)
        dev.off()
        
        #Store resolution vs average silhouette score
        resSil_mat[i,2] = format(round(mean(sil[,3]), 4), nsmall = 4)
      }
      i=i+1
    }
  }
  #Convert character matrix to numeric matrix and assign to appropriate variable
  resSil_mat = matrix(as.numeric(resSil_mat),ncol=ncol(resSil_mat))
  assign(x=paste0("resSil_",objList[k]),value=resSil_mat)
}


```

#### Clustering without batch correction
```{r clusterResultsMerged,fig.width=15, fig.height=9, echo=FALSE, results="asis"}
for (res in resolution){
  resMod = as.numeric(res)
  fileName = paste0(params$imageDir,"/silhouetteResolution_",res,"_merged.pdf")
  cat("  \n##### Clustering at Resolution ", res," \n")
  clusterPlot = DimPlot(merged, group.by=paste0("SLM_SCT_snn_res.",resMod),
                        label=T,repel=T,
    order=sort(as.numeric(as.character(unique(merged[[paste0("SLM_SCT_snn_res.",resMod)]][,1]))),
               decreasing=T))+
    ggtitle(paste("Merged clusters at resolution",res))
  p1 = ggdraw(clusterPlot)
  p2 = ggdraw()+draw_image(fileName)
  print(plot_grid(p1,p2))
  cat("  \n")
}

cat("  \n##### Average silhouette scores vs clustering resolution \n")
plot(resSil_merged,lty=3,pch=1,type="b",
     lwd=2,cex=3,
     main="Average silhouette score vs clustering resolution",
     xlab = "Resolution", ylab="Average Silhouette Score",
     xlim=c(min(resSil_merged[,1])-0.1,max(resSil_merged[,1])+0.1),
     ylim=c(round(min(resSil_merged[,2])-0.005,digits=3),
            round(max(resSil_merged[,2])+0.005,digits=3)))
cat("  \n")
```

#### Clustering after batch correction with CCA Integration
```{r clusterIntegrated, fig.width=15, fig.height=9, echo=FALSE, results="asis"}
for (res in resolution){
  resMod = as.numeric(res)
  fileName = paste0(params$imageDir,"/silhouetteResolution_",res,"_integrated.pdf")
  cat("  \n##### Clustering at Resolution ", res," \n")
  clusterPlot = DimPlot(integrated, group.by=paste0("SLM_int_snn_res.",resMod),
                        label=T,repel=T,
    order=sort(as.numeric(as.character(unique(integrated[[paste0("SLM_int_snn_res.",resMod)]][,1]))),
               decreasing=T))+
    ggtitle(paste("Integrated clusters at resolution",res))
  p1 = ggdraw(clusterPlot)
  p2 = ggdraw()+draw_image(fileName)
  print(plot_grid(p1,p2))
  cat("  \n")
}

cat("  \n##### Average silhouette scores vs clustering resolution \n")
plot(resSil_integrated,lty=3,pch=1,type="b",lwd=2,cex=3,
  main="Average silhouette score vs clustering resolution",
  xlab = "Resolution", ylab="Average Silhouette Score",
  xlim=c(min(resSil_integrated[,1])-0.1,max(resSil_integrated[,1])+0.1),
  ylim=c(round(min(resSil_integrated[,2])-0.005,digits=3),
         round(max(resSil_integrated[,2])+0.005,digits=3)))
cat("  \n")
```


#### Clustering after batch correction with Harmony (by group ID)
```{r clusterHarmonyGroup, fig.width=15, fig.height=9, echo=FALSE, results="asis"}
for (res in resolution){
  resMod = as.numeric(res)
  fileName = paste0(params$imageDir,"/silhouetteResolution_",res,"_harmonyGroup.pdf")
  cat("  \n##### Clustering at Resolution ", res," \n")
  clusterPlot = DimPlot(harmonyGroup, group.by=paste0("SLM_SCT_snn_res.",resMod),
                        label=T,repel=T,
    order=sort(as.numeric(as.character(unique(harmonyGroup[[paste0("SLM_SCT_snn_res.",resMod)]][,1]))),
               decreasing=T))+
    ggtitle(paste("Harmony (group) clusters at resolution",res))
  p1 = ggdraw(clusterPlot)
  p2 = ggdraw()+draw_image(fileName)
  print(plot_grid(p1,p2))
  cat("  \n")
}

cat("  \n##### Average silhouette scores vs clustering resolution \n")
plot(resSil_harmonyGroup,lty=3,pch=1,type="b",lwd=2,cex=3,
  main="Average silhouette score vs clustering resolution",
  xlab = "Resolution", ylab="Average Silhouette Score",
  xlim=c(min(resSil_harmonyGroup[,1])-0.1,max(resSil_harmonyGroup[,1])+0.1),
  ylim=c(round(min(resSil_harmonyGroup[,2])-0.005,digits=3),
         round(max(resSil_harmonyGroup[,2])+0.005,
               digits=3)))

cat("  \n")
```


#### Clustering after batch correction with Harmony (by sample ID)
```{r clusterHarmonySample, fig.width=15, fig.height=9, echo=FALSE, results="asis"}
for (res in resolution){
  resMod = as.numeric(res)
  fileName = paste0(params$imageDir,"/silhouetteResolution_",res,"_harmonySample.pdf")
  cat("  \n##### Clustering at Resolution ", res," \n")
  clusterPlot = DimPlot(harmonySample,
                        group.by=paste0("SLM_SCT_snn_res.",resMod),
                        label=T,repel=T,
    order=sort(as.numeric(as.character(unique(harmonySample[[paste0("SLM_SCT_snn_res.",
                                                                    resMod)]][,1]))),
               decreasing=T))+
    ggtitle(paste("Harmony (sample) clusters at resolution",res))
  p1 = ggdraw(clusterPlot)
  p2 = ggdraw()+draw_image(fileName)
  print(plot_grid(p1,p2))
  cat("  \n")
}

cat("  \n##### Average silhouette scores vs clustering resolution \n")
plot(resSil_harmonySample,lty=3,pch=1,type="b",
     lwd=2,cex=3,
  main="Average silhouette score vs clustering resolution",
  xlab = "Resolution", ylab="Average Silhouette Score",
  xlim=c(min(resSil_harmonySample[,1])-0.1,
         max(resSil_harmonySample[,1])+0.1),
  ylim=c(round(min(resSil_harmonyGroup[,2])-0.005,digits=3),
         round(max(resSil_harmonySample[,2])+0.005,digits=3)))
cat("  \n")
```

#### Clustering after batch correction with Reciprocal PCA
```{r clusterRPCA, fig.width=15, fig.height=9, echo=FALSE, results="asis"}
for (res in resolution){
  resMod = as.numeric(res)
  fileName = paste0(params$imageDir,"/silhouetteResolution_",
                    res,"_rpca.pdf")
  cat("  \n##### Clustering at Resolution ", res," \n")
  clusterPlot = DimPlot(rpca, group.by=paste0("SLM_int_snn_res.",resMod),
                        label=T,repel=T,
    order=sort(as.numeric(as.character(unique(rpca[[paste0("SLM_int_snn_res.",
                                                           resMod)]][,1]))),decreasing=T))+
    ggtitle(paste("RPCA clusters at resolution",res))
  p1 = ggdraw(clusterPlot)
  p2 = ggdraw()+draw_image(fileName)
  print(plot_grid(p1,p2))
  cat("  \n")
}

cat("  \n##### Average silhouette scores vs clustering resolution \n")
plot(resSil_rpca,lty=3,pch=1,type="b",lwd=2,cex=3,
  main="Average silhouette score vs clustering resolution",
  xlab = "Resolution", ylab="Average Silhouette Score",
  xlim=c(min(resSil_rpca[,1])-0.1,max(resSil_rpca[,1])+0.1),
  ylim=c(round(min(resSil_rpca[,2])-0.005,digits=3),round(max(resSil_rpca[,2])+0.005,
                                                          digits=3)))
cat("  \n")
```