---
title: "Sample QC Report"
author: "CCBR scRNASeq Pipeliner"
date: '`r format(Sys.time(), "%a %b %d %Y - %X")`'
output:
  html_document: default
params:
  sample: "sampleName"
  resolution: "csvListResolutionString"
  imageDir: "path/to/outputImageDir"
  ref: "speciesAbbrev"
  outFile: "outFile"
  annot: "annotationRef"

geometry:
  margins:0.25in
---

```{r setup, include=FALSE,fig.align='center'}
require("knitr")
knitr::opts_chunk$set(
  echo=FALSE,
  message=FALSE,
  warning=FALSE
)
knitr::opts_knit$set(root.dir = "../..")
message(getwd())
```

```{r libraryImport, message=FALSE, warning=FALSE}
.libPaths("/data/CCBR_Pipeliner/db/PipeDB")
.libPaths("/data/CCBR_Pipeliner/db/PipeDB/scrna4.2Rlibs")

#Visualization libraries
library(magick)
library(ggplot2)
library(cowplot)
library(scales)
library(RColorBrewer)

#Bioconductor dependencies
library(BiocGenerics)
library(Biobase)
library(farver)
library(S4Vectors)
library(AnnotationDbi)
# library(modes)
library(dplyr)
library(Matrix)
library(reshape2)
library(tools)
library(stringr)

#statistics libraries
# library(Routliers)
library(VennDiagram)
library(cluster)
library(DT)

#scRNASeq libraries
library(Seurat)
library(scRNAseq)
library(SingleCellExperiment)
library(SingleR)
library(scater)
library(DoubletFinder)
# library(URD)
library(celldex)
library(Orthology.eg.db)
library(org.Mm.eg.db)
library(org.Hs.eg.db)
library(flexmix)
library(SeuratWrappers)
library(djvdj)
```

```{r dataImport, message=TRUE, warning=FALSE, include=FALSE}
sampleName = as.character(params$sample)
h5_fileName = paste0("cellRangerOut/",sampleName,"/outs/filtered_feature_bc_matrix.h5")
message(h5_fileName)
species = as.character(params$ref)
outFile = as.character(params$outFile)
resolutionString = as.character(strsplit(gsub(",+",",",params$resolution),split=",")[[1]])
resolution = as.numeric(strsplit(gsub(",+",",",params$resolution),split=",")[[1]])
imageDir = as.character(params$imageDir)
annotDB = as.character(params$annot)
groupFile = read.delim("groups.tab",header=F,stringsAsFactors = F)
sample = groupFile$V3[groupFile$V1 == tail(strsplit(h5_fileName,"/")[[1]],3)[1] & groupFile$V4 == "gex"]

rnaCounts = Read10X_h5(h5_fileName)
if(class(rnaCounts) == "list") {rnaCounts = rnaCounts$'Gene Expression'}

if (species=="both"){
  geneIds = gsub(".*_","",rownames(rnaCounts))
  speciesId = gsub("_.*","",rownames(rnaCounts))
  newIds = paste(geneIds,speciesId,sep="_")
  rownames(rnaCounts)=newIds
}

#Create Seurat object
so <- CreateSeuratObject(rnaCounts)
# groupFile = read.delim("groups.tab",header=F,stringsAsFactors = F)
# sample = groupFile$V3[groupFile$V1 == tail(strsplit(h5_fileName,"/")[[1]],3)[1] & groupFile$V4 == "gex"]
# print(groupFile[groupFile$V3 == sample & groupFile$V4 == "vdj",])
if (nrow(groupFile[groupFile$V3 == sample & groupFile$V4 == "vdj",]) > 0 ) {
  tcrSamples = groupFile$V1[groupFile$V3 == sample & groupFile$V4 == "vdj"]
  tcrSamples = paste0("cellrangerOut/",tcrSamples,"/outs")

  so = import_vdj(input = so, vdj_dir = tcrSamples,  filter_paired = FALSE  )

}

fileTitle=paste("Initial Quality Control:", sample)
```

## `r fileTitle`

```{r functionNames, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
seuratClustering = function(so){
  so$Sample = sample
  ###CHECK CODE HERE FOR PC SELECTION?
  npcs = 30
  so@misc$npcs = npcs

  so <- RunPCA(object = so, features = VariableFeatures(object = so), do.print = TRUE, pcs.print = 1:5,genes.print = 0,verbose=F,npcs = 30)
  so <- FindNeighbors(so,dims = 1:npcs)
  for(res in resolution){
    so <- FindClusters(so, verbose=F, resolution = res, algorithm = 3)
    vect1 = so$seurat_clusters
    names(vect1) = colnames(so)
    so[[paste0("SLM_res_sample.",res)]] = vect1
    so <- FindClusters(so, verbose=F, resolution = res, algorithm = 4)
    vect2 = so$seurat_clusters
    names(vect2) = colnames(vect2)
    so[[paste0("Leiden_res_sample.",res)]] = vect2
  }

  so <- RunUMAP(so,dims = 1:npcs,n.components = 3L)

  return(so)
}

doublets <-function(dfso){
  dfso <- SCTransform(dfso)
  dfso <- RunPCA(dfso, pc.genes = dfso@var.genes, pcs.print = 0,verbose = F,npcs =10)
  npcs = 10
  dfso <- RunUMAP(dfso, verbose=TRUE,dims = 1:npcs)


  sweep.res.list_kidney <- paramSweep_v3(dfso,PCs = 1:10, sct = T)
  sweep.stats_kidney <- summarizeSweep(sweep.res.list_kidney, GT = FALSE)
  # print("aaaaaaaaaaaaaaaaaaaaaaaaaaaaaa")
  bcmvn_kidney <- find.pK(sweep.stats_kidney)
  ## pK Identification (ground-truth) ------------------------------------------------------------------------------------------

  ## Homotypic Doublet Proportion Estimate -------------------------------------------------------------------------------------
  homotypic.prop <- modelHomotypic(dfso$annot)
  perc = 0.005 * (length(colnames(dfso))/1000)
  nExp_poi <- round(perc*length(colnames(dfso)))#dfso@cell.names
  nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))
  # print("xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx")

  ## Run DoubletFinder with varying classification stringencies ----------------------------------------------------------------
  dfso <- doubletFinder_v3(dfso, pN = 0.25, pK = 0.09, nExp = nExp_poi, reuse.pANN = FALSE,PCs = 1:10,sct = T)
  pAAN=tail(names(dfso@meta.data),2)[1]
  dfso <- doubletFinder_v3(dfso, pN = 0.25, pK = 0.09, nExp = nExp_poi.adj, reuse.pANN = pAAN,PCs = 1:10,sct = T)

  return(dfso)
}

convertHumanGeneList <- function(gns){
  egs <- mapIds(org.Hs.eg.db, gns, "ENTREZID","SYMBOL")
  mapped <- AnnotationDbi::select(Orthology.eg.db, egs, "Mus.musculus","Homo.sapiens")
  mapped$MUS <- mapIds(org.Mm.eg.db, as.character(mapped$Mus.musculus), "SYMBOL", "ENTREZID")
  return(as.character(unlist(mapped$MUS )))
}


runSingleR = function(obj,refFile,fineORmain){
  obj = DietSeurat(obj, graphs = "umap")
  sce = as.SingleCellExperiment(obj,assay = "SCT")
  rownames(sce) = gsub("-GRCh38","",rownames(sce))
  rownames(sce) = gsub("-mm10","",rownames(sce))
  ref = refFile
  s = SingleR(test = sce, ref = ref,labels = ref[[fineORmain]])
  return(s$pruned.labels)
  print(head(s$pruned.labels))
}

```

```{r initialQC, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}

if(species=="hg38"){so[["percent.mt"]] <- PercentageFeatureSet(so, pattern = "^MT-")}
if(species=="mm10"){so[["percent.mt"]] <- PercentageFeatureSet(so, pattern = "^mt-")}

if(species=="both"){
  so[["percent.mt.hg38"]] <- PercentageFeatureSet(so, pattern = "^MT-")
  so[["percent.mt.mm10"]] <- PercentageFeatureSet(so, pattern = "^mt-")
  so[["percent.mt"]] = as.numeric(so$percent.mt.hg38) + as.numeric(so$percent.mt.mm10)
}

prefilter_count = ncol(so)

so <- subset(so, subset = nFeature_RNA > 200) #Remove cells with less than 200 unique genes
#Remove cells below the 3 MADs threshold for unique genes and total reads, and above the 3 MADs threshold for mitochondrial percentage
# nCount_out = outliers_mad(log2(so$nCount_RNA),threshold = 3)$LL_CI_MAD
# nFeature_out = outliers_mad(log2(so$nFeature_RNA),threshold = 3)$LL_CI_MAD
# mt_out = outliers_mad(so$percent.mt,threshold = 3)$UL_CI_MAD
#
# cellsToRemove.Feature= colnames(so)[which(log2(so$nFeature_RNA)<nFeature_out)]
# cellsToRemove.Count = colnames(so)[which(log2(so$nCount_RNA)<nCount_out)]
# cellsToRemove.Mito = colnames(so)[which(so$percent.mt>mt_out)]
#
# #Preliminary stats images

# par(mar = c(5, 4, 4, 2) + 0.1)
#
#Venn Diagram of initial filtered cells
# colorList=hue_pal()(3)
# venn.diagram(x=list(cellsToRemove.Feature,cellsToRemove.Count,cellsToRemove.Mito),
# 			 filename=paste0(imageDir,"/cellsRemovedVenn_",sample,".png"),
#        category.names=c("Below feature threshold","Below count threshold","Above mitochondrial threshold"),
# 			 main = paste(sample,"filtered cells:",
# 						  length(unique(c(cellsToRemove.Feature,cellsToRemove.Count,cellsToRemove.Mito),sep=" "))),
# 			 cat.dist=c(-0.07,-0.07,-0.07),
# 			 fill = colorList,
# 			 alpha = 0.5
# )
#
#
#

#Generate initial filter data with MiQC

so <- RunMiQC(so, percent.mt = "percent.mt", nFeature_RNA = "nFeature_RNA", posterior.cutoff = 0.7,model.slot = "flexmix_model")
so_prefilter = so
so = subset(so, miQC.keep == "keep")

if(species=="both"){
  so$nCount_GRCh38 = apply(so@assays$RNA@counts[grep("GRCh38",rownames(so)),],2,sum)
  so$nCount_mm10 = apply(so@assays$RNA@counts[grep("mm10",rownames(so)),],2,sum)
  so$nFeature_GRCh38 = apply(so@assays$RNA@counts[grep("GRCh38",rownames(so)),],2,function(x) length(which(x>0)))
  so$nFeature_mm10 = apply(so@assays$RNA@counts[grep("mm10",rownames(so)),],2, function(x) length(which(x>0)))

  so$countFraction_GRCh38 = so$nCount_GRCh38/so$nCount_RNA
  so$countFraction_mm10 = so$nCount_mm10/so$nCount_RNA
  so$featureFraction_GRCh38 = so$nFeature_GRCh38/so$nFeature_RNA
  so$featureFraction_mm10 = so$nFeature_mm10/so$nFeature_RNA

  so$countID = "human"
  so$countID[which(so$countFraction_mm10>so$countFraction_GRCh38)] = "mouse"
  so$countKeep = "keep"
  so$countKeep[which(apply(cbind(so$countFraction_mm10,so$countFraction_GRCh38),1,min)>0.15)] = "discard"
  so$featureID = "human"
  so$featureID[which(so$featureFraction_mm10>so$featureFraction_GRCh38)] = "mouse"
  so$featureKeep = "keep"
  so$featureKeep[which(apply(cbind(so$featureFraction_mm10,so$featureFraction_GRCh38),1,min)>0.15)]="discard"

  so=subset(so,cells=colnames(so)[which(so$featureKeep=="keep" & so$countKeep=="keep")])
}

#Pre filter violin plots
vln_nFeature_preFilter=ggplot(so_prefilter@meta.data,aes(x=orig.ident,y=log2(as.numeric(as.character(nFeature_RNA))),fill=orig.ident))
vln_nFeature_preFilter = vln_nFeature_preFilter+geom_violin(trim=F)+scale_fill_manual(values=hue_pal()(3)[1],labels=c("nFeature"))+
	geom_boxplot(width=0.1,fill="white")+
	scale_x_discrete(labels=NULL)+
	theme_classic()+#geom_hline(yintercept=nFeature_out,linetype="dashed")+
  labs(x=paste(sample,"Pre-filter"), y="Log2(nFeature_RNA)")+
	labs(fill = paste(sample,"Pre-filter"),title="Pre-filter nFeature_RNA")
vln_nCount_preFilter = ggplot(so_prefilter@meta.data,aes(x=orig.ident,y=log2(as.numeric(as.character(nCount_RNA))),fill=orig.ident))
vln_nCount_preFilter = vln_nCount_preFilter+geom_violin(trim=F)+scale_fill_manual(values=hue_pal()(3)[2],labels=c("nCount"))+
	geom_boxplot(width=0.1,fill="white")+
	scale_x_discrete(labels=NULL)+
	theme_classic()+#geom_hline(yintercept=nCount_out,linetype="dashed")+
  labs(x=paste(sample,"Pre-filter"), y="Log2(nCount_RNA)")+
	labs(fill = paste(sample,"Pre-filter"),title="Pre-filter nCount_RNA")
vln_pctMito_preFilter=ggplot(so_prefilter@meta.data,aes(x=orig.ident,y=log2(as.numeric(as.character(percent.mt))),fill=orig.ident))
vln_pctMito_preFilter=vln_pctMito_preFilter+geom_violin(trim=F)+scale_fill_manual(values=hue_pal()(3)[3],labels=c("Percent Mito"))+
	geom_boxplot(width=0.1,fill="white")+
	scale_x_discrete(labels=NULL)+
	theme_classic()+#geom_hline(yintercept=log2(mt_out),linetype="dashed")+
  labs(x=paste(sample,"Pre-filter"), y="Log2(Percent Mito)")+
	labs(fill = paste(sample,"Pre-filter"),title="Pre-filter Percent Mito")

postfilter_count=ncol(so)

#post filter violin plots
vln_nFeature_postFilter=ggplot(so@meta.data,aes(x=orig.ident,y=log2(as.numeric(as.character(nFeature_RNA))),fill=orig.ident))
vln_nFeature_postFilter = vln_nFeature_postFilter+geom_violin(trim=F)+scale_fill_manual(values=hue_pal()(3)[1],labels=c("nFeature"))+
	geom_boxplot(width=0.1,fill="white")+
	scale_x_discrete(labels=NULL)+
	theme_classic()+#geom_hline(yintercept=log2(nFeature_out),linetype="dashed")+
  labs(x=paste(sample,"Post-filter"), y="Log2(nFeature_RNA)")+
	labs(fill = paste(sample,"Post-filter"),title="Post-filter nFeature_RNA")
vln_nCount_postFilter = ggplot(so@meta.data,aes(x=orig.ident,y=log2(as.numeric(as.character(nCount_RNA))),fill=orig.ident))
vln_nCount_postFilter = vln_nCount_postFilter+geom_violin(trim=F)+scale_fill_manual(values=hue_pal()(3)[2],labels=c("nCount"))+
	geom_boxplot(width=0.1,fill="white")+
	scale_x_discrete(labels=NULL)+
	theme_classic()+#geom_hline(yintercept=nCount_out,linetype="dashed")+
  labs(x=paste(sample,"Post-filter"), y="Log2(nCount_RNA)")+
	labs(fill = paste(sample,"Post-filter"),title="Post-filter nCount_RNA")
vln_pctMito_postFilter=ggplot(so@meta.data,aes(x=orig.ident,y=log2(as.numeric(as.character(percent.mt))),fill=orig.ident))
vln_pctMito_postFilter=vln_pctMito_postFilter+geom_violin(trim=F)+scale_fill_manual(values=hue_pal()(3)[3],labels=c("Percent Mito"))+
	geom_boxplot(width=0.1,fill="white")+
	scale_x_discrete(labels=NULL)+
	theme_classic()+#geom_hline(yintercept=log2(mt_out),linetype="dashed")+
  labs(x=paste(sample,"Post-filter"), y="Log2(Percent Mito)")+
	labs(fill = paste(sample,"Post-filter"),title="Post-filter Percent Mito")


if(species=="hg38"){
  s.genes <- cc.genes$s.genes
  g2m.genes <- cc.genes$g2m.genes
}

if(species=="mm10"){
  s.genes <- convertHumanGeneList(cc.genes$s.genes)
  g2m.genes <- convertHumanGeneList(cc.genes$g2m.genes)
}

if(species=="both"){
  s.genes.GRCh38 <- paste(cc.genes$s.genes,"GRCh38",sep="-")
  g2m.genes.GRCh38 <- paste(cc.genes$g2m.genes,"GRCh38",sep="-")
  s.genes.mm10 <- paste(convertHumanGeneList(cc.genes$s.genes),"mm10",sep="-")
  g2m.genes.mm10 <- paste(convertHumanGeneList(cc.genes$g2m.genes),"mm10",sep="-")

}
#SCTransform normalization and LogNormalize+Scaling on RNA Assay (available if needed for downstream analysis)
so = SCTransform(so)
so = NormalizeData(so, normalization.method = "LogNormalize", scale.factor = 10000,assay = "RNA")
so = ScaleData(so,assay = "RNA")
so_prefilter = SCTransform(so_prefilter)

if(species=="both"){
  so = CellCycleScoring(so,s.features = s.genes.GRCh38, g2m.features = g2m.genes.GRCh38, set.ident = TRUE)
  so$G2M.Score.GRCh38 = so$G2M.Score
  so$S.Score.GRCh38 = so$S.Score
  so$Phase.GRCh38 = so$Phase
  so = CellCycleScoring(so,s.features = s.genes.mm10, g2m.features = g2m.genes.mm10, set.ident = TRUE)
  so$G2M.Score.mm10 = so$G2M.Score
  so$S.Score.mm10 = so$S.Score
  so$Phase.mm10 = so$Phase
} else {
  so = CellCycleScoring(so,s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
  so_prefilter = CellCycleScoring(so_prefilter,s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
}

so = seuratClustering(so)
so_prefilter = seuratClustering(so_prefilter)


if(species == "hg38" | species=="both"){
  so$HPCA_main <- runSingleR(so,celldex::HumanPrimaryCellAtlasData(),"label.main")
  so$HPCA <-  runSingleR(so,celldex::HumanPrimaryCellAtlasData(),"label.fine")
  so$BP_encode_main <-  runSingleR(so,celldex::BlueprintEncodeData(),"label.main")
  so$BP_encode <-  runSingleR(so,celldex::BlueprintEncodeData(),"label.fine")
  so$monaco_main <-  runSingleR(so,celldex::MonacoImmuneData(),"label.main")
  so$monaco <-     runSingleR(so,celldex::MonacoImmuneData(),"label.fine")
  so$immu_cell_exp_main <-  runSingleR(so,celldex::DatabaseImmuneCellExpressionData(),"label.main")
  so$immu_cell_exp <- runSingleR(so,celldex::DatabaseImmuneCellExpressionData(),"label.fine")
  so$annot = so[[paste0(annotDB,"_main")]]
}

if(species == "mm10" | species=="both"){
  so$immgen_main <-  runSingleR(so,celldex::ImmGenData(),"label.main")
  so$immgen <- runSingleR(so,celldex::ImmGenData(),"label.fine")
  so$mouseRNAseq_main <-  runSingleR(so,celldex::MouseRNAseqData(),"label.main")
  so$mouseRNAseq <- runSingleR(so,celldex::MouseRNAseqData(),"label.fine")
  so$annot = so[[paste0(annotDB,"_main")]]
}


dfso = doublets(so)
so$DF_hi.lo = dfso[[tail(names(dfso@meta.data),1)]]

##doublet DimPlot
doubletCount = length(which(so$DF_hi.lo =="Doublet"))
# print(doubletPlot+labs(title = paste0(sample," Doublets")))
so_prefilter$DF_hi.lo = so$DF_hi.lo
so_prefilter$annot = so$annot
so_prefilter$annot[setdiff(colnames(so_prefilter),colnames(so))] <- "Filtered Cells"
so=subset(so,cells=names(so$DF_hi.lo)[so$DF_hi.lo =="Singlet"])

filtered = vector(length=ncol(so_prefilter))
filtered[] = "Retained"
filtered[which(so_prefilter$miQC.keep=="discard")] = "Filtered: MiQC"
filtered[which(so_prefilter$DF_hi.lo == "Doublet")] = "Filtered: Doublets"
so_prefilter$filtered = filtered

saveRDS(so,outFile)

```

```{r qcReportImages, message = FALSE, warning = FALSE, echo=FALSE, include=FALSE}
#Preliminary stats table : need to rework into sortable table
png(paste0(imageDir,"/filterStats_",sample,".png"),height=3000,width=3000,res=500,units="px")
par(mar=c(0,0,0,0))
plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
text(x=0.05, y=0.95, paste0("QC Metrics for ",sample,":"),
     cex = 1.5, col = "black", family="serif", font=2, adj=0)

text(x=0.075, y=0.9, paste0("Cells before filtering: ", prefilter_count),
     cex = 1.25, col = "black", family="serif", font=3, adj=0)
text(x=0.075, y=0.865, paste0("Cells after filtering: ", ncol(so)),
     cex = 1.25, col = "black", family="serif", font=3, adj=0)

text(x=0.1, y=0.815,paste0("Cells removed by MiQC: ",length(which(so_prefilter$miQC.keep == "discard"))),
		 cex = 1, col="dimgray",family="serif", font = 1, adj =0)
text(x=0.1, y=0.725,paste0("Cells removed as doublets: ",doubletCount),
		 cex = 1, col="dimgray",family="serif", font = 1, adj =0)

text(x=0.1, y=0.68, paste0("Number of principal components used: ",so@misc$npcs),
		 cex = 1, col="dimgray",family="serif", font = 1, adj =0)

dev.off()

filterConditionPlot=DimPlot(so_prefilter,group.by="filtered") + ggtitle(paste("Sample",sample,"before filtering"))
filterConditionSplitPlot = DimPlot(so_prefilter,group.by="filtered",split.by="filtered") + ggtitle("Filtered cells in isolation")

miQC_feature_mitoPlot=FeatureScatter(so_prefilter,feature1="nFeature_RNA",feature2="percent.mt",group.by="Sample") + ggtitle("","Number of genes vs mitochondrial percentage") +NoLegend()
miQC_count_mitoPlot=FeatureScatter(so_prefilter,feature1="nCount_RNA",feature2="percent.mt",group.by="Sample") + ggtitle("","Number of reads vs mitochondrial percentage") + NoLegend()
miQC_feature_countPlot=FeatureScatter(so_prefilter,feature1="nFeature_RNA",feature2="nCount_RNA",group.by="Sample") + ggtitle("","Number of genes vs number of counts") + NoLegend()
miQC_probabilityPlot = PlotMiQC(so_prefilter,color.by="miQC.probability") + ggplot2::scale_color_gradient(low = "grey", high = "purple") + ggtitle("MiQC probability of low quality")
miQC_retentionPlot = PlotMiQC(so_prefilter,color.by="miQC.keep") + ggtitle("MiQC classification of retained cells")


cellCyclePlot=DimPlot(so,group.by="Phase")+labs(title = paste0(sample," Cell Cycle"))
singleRPlot = DimPlot(so,group.by="annot",label=T,repel=T) + NoLegend()+ labs(title = paste0(sample," annotations by ",annotDB)) + theme(plot.title = element_text(hjust=0.5))


#cluster analysis
coord = Embeddings(so,reduction='pca')[,1:30]#based on npcs above
objDist = dist(coord, method = "euclidean")
resSil_SLM = matrix(ncol =2, nrow = length(resolutionString))
resSil_SLM[] = NA
resSil_Leiden=resSil_SLM

for (algorithm in c("SLM","Leiden")){
  # if(algorithm == "Leiden"){next}
  for (i in 1:length(resolution)){

    Idents(so) = paste0(algorithm,"_res_sample.",resolution[i])
    ## print cluster DimPlots
    objName = paste("clusterPlot",algorithm,resolution[i],sep="_")
    clusterPlot=DimPlot(so,label=T,repel=T)
    assign(objName,clusterPlot)
    ## print cluster silhouette plot
    clusters = Idents(so)
    pdf(paste0(imageDir,"/silhouetteResolution_",algorithm,"_",resolution[i],".pdf"))
    if(length(unique(clusters))==1){ #edge case of failed sample, only one cluster
  		plot.new()
  		text(0.5,0.5,"No silhouette available for single cluster samples")
  		dev.off()
  		next
  	}
    if(ncol(so) > 90000){
      plot.new()
      text(0.5,0.5,"No silhouette available for datasets with more than 90,000 cells")
      dev.off()
      next
    }
    sil = silhouette(as.numeric(as.character(clusters)),dist = objDist)
    palette=alpha(colour=hue_pal()(length(unique(Idents(so)))),alpha=0.7)
    print(plot(sil, col=palette[as.factor(clusters[order(as.numeric(as.character(clusters)),decreasing=F)])],
  	main=paste0("Silhouette plot of clustering resolution ", resolutionString[i]), lty=2,
  	sub=paste("Average silhouette width:",format(round(mean(sil[,3]), 4), nsmall = 4))))

  	abline(v=mean(sil[,3]), col="red4", lty=2)
  	dev.off()

    if(algorithm=="SLM"){
      resSil_SLM[i,] = c(resolution[i], format(round(mean(sil[,3]), 4), nsmall = 4))
    }else if(algorithm=="Leiden"){
      resSil_Leiden[i,] = c(resolution[i], format(round(mean(sil[,3]), 4), nsmall = 4))
    }

  }
}

```

### Cell Filtering {.tabset}

Initial filtering is performed using the MiQC wrapper for Seurat (http://htmlpreview.github.io/?https://github.com/satijalab/seurat-wrappers/blob/master/docs/miQC.html).
MiQC uses a mixture model to identify low quality cells based on the number of detected genes (`nFeature_RNA`) and the mitochondrial percentage (`percent.mt`) and assigns
a probability that the cell in question should be discarded prior to analysis.


```{r prelimStats, fig.width=15, fig.height=7}
p1 = ggdraw() + draw_image(list.files(path=imageDir,pattern="filterStats_",full.names=T))
p2 = ggdraw(filterConditionPlot)
p3 = ggdraw(DimPlot(so,group.by="Sample")+ggtitle("Post-filtering"))
p4 = ggdraw(filterConditionSplitPlot)

p1 #statistics
plot_grid(p2, p3,ncol=2) #Pre-filter and post-filter images
p4 #split image
##code for table of statistics here

```

#### Filtering with MiQC
```{r miQC, fig.width=15,fig.height=7}
p1=ggdraw(miQC_feature_mitoPlot)
p2=ggdraw(miQC_count_mitoPlot)
p3 = ggdraw(miQC_feature_countPlot)

plot_grid(p1,p2,p3,ncol=3)

p4=ggdraw(miQC_probabilityPlot)
p5=ggdraw(miQC_retentionPlot)
plot_grid(p4,p5,ncol=2)

```


#### Violin Plots

```{r violinPlots, fig.width=15, fig.height=12, echo=FALSE, results="asis"}
p1 = ggdraw(vln_nFeature_preFilter)
p2 = ggdraw(vln_nFeature_postFilter)

p3 = ggdraw(vln_nCount_preFilter)
p4 = ggdraw(vln_nCount_postFilter)

p5 = ggdraw(vln_pctMito_preFilter)
p6 = ggdraw(vln_pctMito_postFilter)
plot_grid(p1,p3,p5,p2,p4,p6, ncol=3)
```


### Cell Annotation {.tabset}

#### Cell Cycle
```{r CellCycle, fig.width=6, fig.height=6}
p1<- ggdraw(cellCyclePlot)
plot_grid(p1)
```

#### SingleR
```{r SingleR, fig.width=6, fig.height=6}
p1 <- ggdraw(singleRPlot)
plot_grid(p1)
datatable(as.data.frame.matrix(table(so$annot,so$Sample)),
  colnames=c("nCells"))

```

### Initial clustering {.tabset}

#### Clustering with the SLM algorithm

```{r clusteringResults_SLM,fig.width=15,fig.height=9,echo=FALSE,results="asis"}
for (res in resolutionString){
  resMod = as.numeric(res)
  fileName = paste0(imageDir,"/silhouetteResolution_SLM_",resMod,".pdf")
  cat("  \n##### Clustering at Resolution ", res," \n")
  clusterPlot = DimPlot(so, group.by=paste0("SLM_res_sample.",resMod),label=T,repel=T,
    order=sort(as.numeric(as.character(unique(so[[paste0("SLM_res_sample.",resMod)]][,1]))),decreasing=T))+
    ggtitle(paste("SLM clustering at resolution",res))
  p1 = ggdraw(clusterPlot)
  p2 = ggdraw()+draw_image(fileName)
  print(plot_grid(p1,p2))
  cat("  \n")
}

cat("  \n##### Average silhouette scores vs clustering resolution \n")
plot(resSil_SLM,lty=3,pch=1,type="b",lwd=2,cex=3,
  main="Average silhouette score vs clustering resolution",xlab = "Resolution", ylab="Average Silhouette Score",
  xlim=c(min(as.numeric(resSil_SLM[,1]))-0.1,max(as.numeric(resSil_SLM[,1]))+0.1),
  ylim=c(round(min(as.numeric(resSil_SLM[,2]))-0.005,digits=3),round(max(as.numeric(resSil_SLM[,2]))+0.005,digits=3)))

cat("  \n")
```


#### Clustering with the Leiden algorithm

```{r clusteringResults_Leiden,fig.width=15,fig.height=9,echo=FALSE,results="asis"}
for (res in resolutionString){
  resMod = as.numeric(res)
  fileName = paste0(imageDir,"/silhouetteResolution_Leiden_",resMod,".pdf")
  cat("  \n##### Clustering at Resolution ", res," \n")
  clusterPlot = DimPlot(so, group.by=paste0("Leiden_res_sample.",resMod),label=T,repel=T,
    order=sort(as.numeric(as.character(unique(so[[paste0("Leiden_res_sample.",resMod)]][,1]))),decreasing=T))+
    ggtitle(paste("Leiden clustering at resolution",res))
  p1 = ggdraw(clusterPlot)
  p2 = ggdraw()+draw_image(fileName)
  print(plot_grid(p1,p2))
  cat("  \n")
}


cat("  \n##### Average silhouette scores vs clustering resolution \n")
plot(resSil_Leiden,lty=3,pch=1,type="b",lwd=2,cex=3,
  main="Average silhouette score vs clustering resolution",xlab = "Resolution", ylab="Average Silhouette Score",
  xlim=c(min(as.numeric(resSil_Leiden[,1]))-0.1,max(as.numeric(resSil_Leiden[,1]))+0.1),
  ylim=c(round(min(as.numeric(resSil_Leiden[,2]))-0.005,digits=3),round(max(as.numeric(resSil_Leiden[,2]))+0.005,digits=3)))

cat("  \n")
```


### Session Info
```{r}
sessionInfo()
```
